ESBMC-GPU \\ A Context-Bounded Model Checking Tool \\ to Verify CUDA Programs

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
pointing out changes in manuscript
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

A distinct feature of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC)~\cite{Biere:2009} allied to Satisfiability Modulo Theories (SMT)~\cite{Barrett:2009}, with explicit state-space exploration~\cite{cordeiro:2011,cordeiro:2012}. In summary, concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled; however, even with BMC, state-space exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}. As a consequence, redundant interleavings are eliminated, without ignoring a program's behavior.

@@

ESBMC-GPU is built on top of ESBMC, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.

@@

\noindent \textbf{1. CUDA Operational Model.} An operational model for CUDA libraries that provides support to CUDA functionalities, in conjunction with ESBMC, as shown in Fig.~\ref{figure:arch}. Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, consists of an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf. Sec.~\ref{sec:features}). Indeed, COM allows the necessary control for performing code analysis, where both CUDA operation and knowledge for model checking its properties are available.

@@

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also reduce the state space, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.% \textcolor{red}{similares? entao, qual eh o impacto de se verificar apenas um? eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.

@@

In order to check the aforementioned properties, ESBMC-GPU explicitly explores the possible interleavings (up to the given context bound) and calls the single-threaded BMC procedure on each one, whenever it reaches an RT leaf node. Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings have been systematically explored~\cite{pereira:2016}.

@@

{\bf Performance.} MPOR resulted in a performance improvement of approximately $80\%$, by decreasing the verification time from $16$ to $3$ hours, while the two-threads analysis further reduced that to $789.6$ sec. Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools: GPUVerify ($98.36$ sec), GKLEE ($108.32$ sec), and CIVL ($708.52$ sec). This is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}. %In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications. 
On the other hand, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.

>>>>

A distinct feature of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC)~\cite{Biere:2009} allied to Satisfiability Modulo Theories (SMT)~\cite{Barrett:2009}, with explicit \textcolor{blue}{state-space exploration}~\cite{cordeiro:2011,cordeiro:2012}. In summary, concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled; however, even with BMC, state-space exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}. As a consequence, redundant interleavings are eliminated, without ignoring a program's behavior.

@@

\textcolor{blue}{ESBMC-GPU is built on top of ESBMC}, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.

@@

\noindent \textbf{1. CUDA Operational Model.} An operational model for CUDA libraries that provides support to CUDA functionalities, in conjunction with ESBMC, as shown in Fig.~\ref{figure:arch}. Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, \textcolor{blue}{consists of} an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf. Sec.~\ref{sec:features}). Indeed, COM allows the necessary control for performing code analysis, where both CUDA operation and knowledge for model checking its properties are available.

@@

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also \textcolor{blue}{reduce the state space}, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.% \textcolor{red}{similares? entao, qual eh o impacto de se verificar apenas um? eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.

@@

In order to check the aforementioned properties, ESBMC-GPU explicitly explores the possible interleavings (up to the given context bound) and calls the single-threaded BMC procedure on each one, whenever it reaches an RT leaf node. Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings \textcolor{blue}{have been systematically explored}~\cite{pereira:2016}.

@@

{\bf Performance.} MPOR resulted in a performance improvement of approximately $80\%$, by decreasing the verification time from $16$ to $3$ hours, while the two-threads analysis further reduced that to $789.6$ sec. Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools: GPUVerify ($98.36$ sec), GKLEE ($108.32$ sec), and CIVL ($708.52$ sec). This is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}. %In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications. 
\textcolor{blue}{On the other hand}, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
